### 2017-06-19 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
pushing updates

---
title: "MSDS 7333 Case Study 06"
author: "Matthew Baldree, Ben Brock, Tom Elkins, Austin Kelly"
date: "June 17, 2017"
output:
  html_notebook: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assignment: 
1. Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both: 
    + chosenAP = "00:0f:a3:39:e1:c0"
    + rejectedAP = "00:0f:a3:39:dd:cd"
* *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
* *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*
* *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*

***

2. Implement alternative k-nearest precition method using weights on received signal strength.
* *For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?*
* *Use calcError() to compare approaches*
<font color = "green">

## Introduction
#### *Explain the case study*
- Discuss the goal of the case study

For this case study, the overall goal is to use K-Nearest Neighbors (kNN) and an alternative weighted method to be able to determine how different handheld devices moved through a building <font color = "red">(at Dartmouth. Let's confirm this and add some more detail)</font>. This can be achieved with an analysis of how the different Media Access Control (MAC) addresses for each of the devices interacted with the Access Points (APs) of the building. In this case, the APs are routers with static locations throughout the building.

- Discuss the different Access points and MAC Addresses

The access points as stated above allow for the tracking of the devices thanks to the logged history of times, signal strengths and angles for each MAC address with respect to each access point.<font color = "red">&larr;(consider fragmenting this sentence)</font> From the information that has been given, there are known to be six APs on the building floor in question. However, this is a problem since there are seven APs recorded. To determine which AP to exclude from the analysis, we will need to compare these two addresses:

    chosenAP   = "00:0f:a3:39:e1:c0"
    rejectedAP = "00:0f:a3:39:dd:cd"

and decide which to exclude and why.

  - Discuss why we're trying to remove one

Removing one of these APs will allow us to remove noise from the resulting model. Since only six APs are recorded as existing on the floor, the presence of a seventh points to external interference. Once the superfluous address has been identified, we will execute testing to determine whether there will be any benefit added to the model by keeping or excluding said data. This decision will ultimately lead to a better prediction of location of each device within this floorplan.

- BRIEF overview of K-Nearest Neighbors

K-Nearest Neighbors (kNN) is an ideal method to use in this case as the goal is to be able to predict the path of each device on their journey through the building. To utilize this method, we use separate testing and training data. The training data are known observations from well documented locations of both the handheld devices and APs. These locations are represented by the angle at which the AP received the signal from the device as well as the overall signal strength received.

With this training data, each observation is evaluated along with the observations which occur in the general vicinity in the training data. These occurrences allow for the different probabilities of occurrence to be proposed through the model. Once this model has been generated for k neighbors, it is then tested against the unobserved test data and adapted to suit what actually occurs. This allows for an accurate gauge of accuracy of the model and is recreated for a span of possible k neighbors to exhaust the total possible number of observations.

- Discuss the alternative weighted method.

In contrast to the kNN method, another option to explore is a weighted method based on the signal strength received at the access point. This method involves defining a new metric where the weight of each observation is inversely proportional to the distance from the test observation. From a topographical standpoint, the observations which are closest to the observation will have a much higher influence than those which exist further away.

- Discuss potential for prediction and what we expect to come of this.

The evaluation of this data will allow us to determine which data will be critical to the creation of an accurate model and which will not. Since we have six access points and seven addresses, this thorough investigation will help us decide which address is simply an external interference and if it is beneficial to keep involving said data. As a result, we expect to have a more thorough understanding of the difference between k-Nearest Neighbors and the weighted alternative.

</font>

***

## The Approach (Explained)

### Part 1

Before we begin, we will need to set the parameters for printing and formatting
```{r}
options(digits = 2)
options(error = recover, warn = 2)
```
Before continuing, we will need to address another file which houses the predefined functions. This helps to save notebook real estate and prevents confusion:
```{r}
source("MSDS7333-baldree-case6-fx.r", print.eval = TRUE)
```

Read offline data
```{r}
offline = readData()
```

Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both:

    * chosenAP = "00:0f:a3:39:e1:c0"
    * rejectedAP = "00:0f:a3:39:dd:cd"

Since we have the two MAC addresses in question, we will declare these addresses as variables to simplify later code.

```{r}
chosenAP = "00:0f:a3:39:e1:c0"
rejectedAP = "00:0f:a3:39:dd:cd"
remainingAP = unique(offline[!offline$mac %in% c(chosenAP, rejectedAP),]$mac)
```

* *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*

To fully answer this question, we will need to explore exactly what is happening with each MAC address. The results of this exploration will define the parameters needed for such a decision.

#### Here we will create a Signal Collection Map
For this exercise, we will plot both the count and locations of the signals recorded for the access points. Even though it would be ideal to plot the access points(APs) directly on the map, we do not know their X and Y coordinates.

```{r}
plotSignalMaps(c(chosenAP, rejectedAP))
```

At first glance, we can see that the maps are quite similar. Unfortunately, the information gained from a visual comparison of these two diagrams does not give us a significant insight into which MAC address should be kept and which should be rejected. The next step to reach our goal will be to look at the individual Signal Strength Distributions

***

#### Signal Strength Distributions
Plotting the individual signal strength distributions for each AP angle will help add perspective to the information we are working with. For the sake of consistency, we will use the same stationary coordinates listed in the course book (coord : (2, 12)) <font color = "red">(Add reference here.)</font>

```{r}
df = subset(offline, posX ==2 & posY == 12 & !(mac %in% remainingAP), c(signal, angle, mac))
plotBoxplotSignalStrength(df)
```

From the literature, we know the higher the decibel-milliwatts means a stronger signal strength. A stronger signal strength indicates a closer physical position to the AP since signal strengths decrease exponentially with distance.<font color="red">Find this in the book and cite it.</font> Comparing the signal strength distributions in this chart, we see "e1:c0" has a stronger signal strength at this location than "dd:cd."

What if we compared the two addresses from the opposite side of the building? 
```{r}
# Let's focus on the fixed point (33,3)
df = subset(offline, posX == 33 & posY == 3 & !(mac %in% remainingAP), c(signal, angle, mac))
plotBoxplotSignalStrength(df)
```
Through the use of another perspective, the MAC address with the weaker signal at (2,12) still has a relatively weak signal when viewed from an AP at the other side of the building. The evidence against the address with the lower signal strength is mounting.

***

#### Time to investigate Signal Density
To properly investigate the difference between these two addresses, the signal strength densities for each address should be plotted for each 45 degree angle. This will help display the defining characteristics for each address. The location used for this analysis is the (24,4) from the book.

```{r, fig.height=10}
df = subset(offline, posX == 24 & posY == 4 & !(mac %in% remainingAP), c(signal, angle, mac))
plotDensitySignalStrength(df)
```

From this plot, it can be seen that the rejectedAP has a higher signal density at lower strength while the acceptedAP has a higher density coverage at a higher strength. Judging by the distributions themselves, it seems as if many of the plots are normally distributed while some of the other plots (i.e. "dd:cd" at 180 and "e1:c0" at 315) are heavily skewed or even bimodal. Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle. This would suggest the median of each distribution varies by angle and attributes to a difference in perceived strength.

This evidence would suggest both addresses are kept within the analysis to increase the coverage for the entire spectrum of strengths.

#### Average Signal Strength Distributions
To examine the distribution for all locations and angles of the two APs in question, we will generate the summary statistics. The summary statistics for all possible combinations of locations and orientations for each AP through the use of a new factor.

```{r}
offlineSummary = createOfflineSummary(offline)
```

### <font color = "red"> &darr; this section may no longer be needed. Now missing from the original code. &darr;</font>  
```{r}
#offline$posXY = paste(offline$posX, offline$posY, sep="-")
#
# create data frames for each combination
#byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function(x) x))
```

Now, we calculate the summary statistics:
```{r}
#signalSummary = lapply(byLocAngleAP, function(oneLoc) {
#  ans = oneLoc[1, ]
#  ans$medSignal = median(oneLoc$signal)
#  ans$avgSignal = mean(oneLoc$signal)
#  ans$num = length(oneLoc$signal)
#  ans$sdSignal = sd(oneLoc$signal)
#  ans$iqrSignal = IQR(oneLoc$signal)
#  ans
#})
#offlineSummary = do.call("rbind", signalSummary)
```
### <font color = "red"> &uarr; this section may no longer be needed. Now missing from the original code. &uarr;</font>

The Standard Deviation of the average signal strength per targeted AP:

```{r}
df = subset(offlineSummary, mac %in% c(chosenAP, rejectedAP), c(sdSignal, avgSignal, mac))
plotStdDevSignalStrength(df)
```

From this figure, it is apparent that there is considerable overlap between the two distributions of signal strength. This is evidence in favor of keeping both addresses in the analysis to provide greater coverage.

#### Let's dive into the K-Nearest Neighbors:
<font color = "red"><b>Discuss KNN here</b></font>

```{r}
macs = unique(offlineSummary$mac)
macs
online = readData("online.final.trace.txt", subMacs = macs)
online$posXY = paste(online$posX, online$posY, sep = "-")
length(unique(online$posXY))
tabonlineXYA = table(online$posXY, online$angle)
#output shows measurements were taken at all angles throughout the floor
tabonlineXYA
```

Organize the data where each AP is in a column.

```{r}
onlineSummary = castOnline(online)
#confirm rejected AP is included
dim(onlineSummary)
names(onlineSummary)
```

<font color = "red">(Continued from baldree-case6.r)</font>

> for k-NN, we will include training data with angles close to point in question since angle matters.
> if we want one angle, then include angles that match the rounded orientation of new observation.
> if we want two angles, then pick two multiples of 45 degrees that flank the new observation's orientation.
> if we want three angles, then pick the closest 45 degree increment and on either side of it.
> what if we didn't want to collapse the signal strengths across the m angles, and
> instead return a set of mx166 signals for each access point?

```{r}
source("MSDS7333-baldree-case6-fx.r", print.eval = TRUE)
```

```{r}
#Iterate through 1 to 3 neighbors with 1 to 3 angles for each scenario to determine best combination
#for cross validation
for (ap in c("None", rejectedAP, chosenAP)){
  if (ap == "None"){
    test = onlineSummary
    train = offlineSummary
    newSignals = test[ , 6:12]
  } else {
    test = onlineSummary[, !(names(onlineSummary) %in% ap)]
    train = subset(offlineSummary, mac != ap)
    newSignals = test[,6:11]
  }

  estXY = predXY(newSignals = newSignals,
                 newAngles = test[, 4],
                 train, 
                 numAngles = 3, k = 3)
  actXY = test[ , c("posX", "posY")]
  err = calcError(estXY, actXY)

  print(paste(ap, err))
}
```

<font color = "green" size = "5px"><b>This is where we will perform cross validation and plot for desired AP for a range of k's </b></font>

Perform cross validation and plot for desired AP for a range of k's for each fold. We want to find k-NN estimates from 1...k and aggregate errors over the v folds.

Choose a v-fold of 11
```{r}
v = 11
```

```{r}

# number of locations
permuteLocs = sample(unique(offlineSummary$posXY))

```

```{r}
# matrix locations into folds
permuteLocs = matrix(permuteLocs, ncol = v, nrow = floor(length(permuteLocs)/v))
```
<font color = "Red"><b>&uarr; come back to this one &uarr;</b></font>

```{r}
# summarize and format offline
keepVars = c("posXY", "posX", "posY", "orientation", "angle")
onlineCVSummary = reshapeSS(offline, keepVars = keepVars, sampleAngle = TRUE)
```

```{r}
# number of k's
K = 10
# errors array
err = rep(0, K)

# loop through folds
for (j in 1:v) {
  testFold = subset(onlineCVSummary, posXY %in% permuteLocs[ , j])
  trainFold = subset(offlineSummary, posXY %in% permuteLocs[ , -j])
  actFold = testFold[ , c("posX", "posY")]

  # loop through neighbors
  for (k in 1:K) {
    estFold = predXY(newSignals = testFold[ , 6:12],
                     newAngles = testFold[ , 4],
                     trainFold, numAngles = 3, k = k)
    err[k] = err[k] + calcError(estFold, actFold)
  }
}
# plot k to sum of squared errors
plotSSErrors(err, K)
```

```{r}
# experiment with a few k's based on above analysis to determine k
# that results in lowest sum squared error on test data.
estXY = predXY(newSignals = onlineSummary[ , 6:12],
               newAngles = onlineSummary[ , 4],
               offlineSummary, numAngles = 3, k = 5)
actXY = onlineSummary[ , c("posX", "posY")]
calcError(estXY, actXY)
```

*** 
#Part 2: Alternatvie k-nearest neighbors
```{r}
# create weights for neighbors based on signal strength and multiple times X and Y and sum neighbors
# to get X and Y.
estXY = predXY(newSignals = onlineSummary[ , 6:12],
               newAngles = onlineSummary[ , 4],
               offlineSummary, numAngles = 3, k = 5, weighted=TRUE)
actXY = onlineSummary[ , c("posX", "posY")]
calcError(estXY, actXY)
```

***